\documentclass[12pt,a4paper]{article}

\input{../settings}

\begin{document}

\input{../header}


\begin{center}
{\bf \Large Sheet 1}

{\large Topics: Octave}

Submission deadline: Oct. 28, 2013\\
Submit to: \texttt{robotmappingtutors@informatik.uni-freiburg.de}
\end{center}

\subsubsection*{General Notice}

The exercises should be solved in groups of two students. In general,
assignments will be published on Monday and should be submitted until
the deadline. The source code of programming exercises should be
submitted via email.

We will be using Octave for the programming exercises. Octave is a
command line program for solving numerical computations. Octave is
mostly compatible with MATLAB and is freely available from
\url{www.octave.org}. It is available for Linux, Mac OS, and Windows.
Install Octave on your system in order to solve the programming
assignments. A quick guide to Octave is given in the Octave cheat sheet
which is available on the website of this lecture.

\subsubsection*{Exercise 1: Getting familiar with Octave}

The purpose of this exercise is to familiarize yourself with Octave and
learn basic commands and operations that you will need when solving the
programming exercises throughout this course.

Go through the provided Octave cheat sheet and try out the different
commands. Ask for help whenever you need it. As pointed out in the
sheet, a very useful Octave command is \texttt{help}. Use it to get
information about the correct way to call any Octave function.

\subsubsection*{Exercise 2: Implementing an odometry model}

Implement an Octave function to compute the pose of a robot based on
given odometry commands and its previous pose. Do not consider the
motion noise here.
%As you saw in the lecture, the first step of a Bayes Filter is
%typically predicting the current pose of a robot according to a certain
%motion model. Your task is thus to implement a simplified version of
%the odomtery model you saw in the lecture (ignoring noise for now).

For this exercise, we provide you with a small Octave framework for
reading log files and to visualize results.  To use it, call the
\texttt{main.m} script. This starts the main loop that computes the pose
of the robot at each time step and plots it in the map. Inside the loop,
the function \texttt{motion\_command} is called to compute the pose of
the robot. Implement the missing parts in the file
\texttt{motion\_command.m} to compute the pose $\vec x_t$ given $\vec
x_{t-1}$ and the odometry command $\vec u_t$. These vectors are in the
following form:

\begin{equation*}
      \vec x_t = \left(\begin{array}{c} x \\ y \\ \theta \end{array}\right) \quad
      \vec u_t = \left(\begin{array}{c} \delta_{rot1} \\ \delta_{trans} \\ \delta_{rot2} \end{array}\right), \quad
\end{equation*}

where $\delta_{rot1}$ is the first rotation command, $\delta_{trans}$ is
the translation command, and $\delta_{rot2}$ is the second rotation
command. The pose is represented by the $3\times1$ vector $x$ in
\texttt{motion\_model.m}. The odometry values can be accessed from the
struct $u$ using $u.r1$, $u.t$, and $u.r2$ respectively.

Compute the new robot pose according to the following motion model:

\begin{equation*}
     \begin{array}{l}
     x_t = x_{t-1} + \delta_{trans} \cos(\theta_{t-1} + \delta_{rot1}) \\
     y_t = y_{t-1} + \delta_{trans} \sin(\theta_{t-1} + \delta_{rot1}) \\
     \theta_t = \theta_{t-1} + \delta_{rot1} + \delta_{rot2}
     \end{array}
\end{equation*}

Test your implementation by running the \texttt{main.m} script. The
script will generate a plot of the new robot pose at each time step and
save an image of it in the \texttt{plots} directory.  While debugging,
run the program only for a few steps by replacing the for-loop in
\texttt{main.m} by something along the lines of \texttt{for t = 1:20}.
You can generate an animation from the saved images using \emph{ffmpeg}
or \emph{mencoder}.  With \emph{ffmpeg} you can use the following
command to generate the animation from inside the \texttt{plots}
directory:

\texttt{ffmpeg -r 10 -b 500000 -i odom\_\%03d.png odom.mp4}

\subsubsection*{Exercise 3: Homogeneous transformations}

A robot pose in a given frame is compactly represented as:
\begin{equation*}
      \vec p = \left(\begin{array}{c} x \\ y \\ \theta \end{array}\right) \quad
\end{equation*}


Alternativley, homogeneous coordinates can be used. This simplifies transformations as they are modeled  in matrix form:
\begin{equation*}
  \vec M =
    \begin{pmatrix}
       \vspace{0.5em} \boldsymbol{R} & \boldsymbol{t}\\
       \boldsymbol{0} & 1\\
    \end{pmatrix}
    =
    \begin{pmatrix}
      \cos{\theta} & -\sin{\theta} & x\\
      \sin{\theta} & \cos{\theta} & y\\
      0 & 0 & 1\\
    \end{pmatrix}
\end{equation*}

% This represents a relative motion of the robot with respect to a given frame.

(a) Implement two Octave funtions:
\begin{itemize}
  \item \textbf{v2t} - takes as input the vector form of the robot pose and outputs the corresponding homogeneous transformation
  \item \textbf{t2v} - takes as input an homogeneous transformation representing the robot pose in the 2D space and outputs the corresponding compact vector
\end{itemize}

Test your implementation chaining four different transformations of your choice.

\vspace{2em}

(b) Given two robot poses $p_1=(x_1, y_1, \theta_1)^T$ and $p_2=(x_2, y_2, \theta_2)^T$, how do you get the relative transformation from $p_1$ to $p_2$?

\vspace{2em}
(c) A 2D point (that we will often call ``landmark'' or ``point of interest'') in the plane can be expressed by its $x$ and $y$ values. In order to get a $3 \times 1$ vector and to be able to make multiplications with the $3 \times 3$ matrices representing the robot poses, we need to express the landmarks in homogeneous coordinates, that in this case means putting the scaling value to 1.
\begin{equation*}
      \vec {poi} = \left(\begin{array}{c} x \\ y \\ 1 \end{array}\right) \quad
\end{equation*}

\vspace{1em}

Given a robot pose $p1$ and an $<x,y>$ observation $z$ of a landmark relative to $p1$.

\begin{equation*}
      \vec x_t = \left(\begin{array}{c} 1 \\ 1 \\ \pi/2 \end{array}\right) \quad
      \vec z = \left(\begin{array}{c} 2 \\ 0  \end{array}\right), \quad
\end{equation*}

\vspace{1em}
Compute the location of the  landmark.

\end{document}
